Weekday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population

### 2023-12-11 manuscript/paper.Rmd

```{r tabfour, echo=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)

@@

\caption{\label{tab:tabmodel}Summary model performance measures.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & $\text{df}^a$ & $\text{AIC}^b$ & $\text{BIC}^c$ & $\text{MSE}^d$ & $R^2$ & Adjusted $R^2$\\
\midrule
Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
\addlinespace
Model 6 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
Model 7 & 41 & 15888.38 & 16117.55 & 173.69  & 0.36 & 0.34\\
Model 8 & 42 & 15849.19 & 16083.94 & 170.11  & 0.37 & 0.36\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
\end{tabular}
\end{table}
```

Likelihood ratio tests showed a statistically significant difference between Model 1 and 2 with evidence of in favor of the latter ($\chi^2_{(1)} = 7064; p>0.05$). Model 2 which adjusted for all potentially confounding covariates in \@ref(tab:tabthree) including an interaction between cholestrol level and age was then compared to Model 3 which adjusted for only covariates we found significant in univariate analyses. Results here similarly showed evidence in favor of Model 2 ($\chi^2_{(1)} = 7979; p>0.05$). Comparing Model 2 to Model 3 revealed a non-statistically difference in model fit suggesting that Model 3, even though smaller performed comparably well. Model 3 was thus selected and compared to Model 4, 5, 6, and 7 with evidence based on the LRT favoring its choice as the best model ($p>0.05$). Additional regression diagnostics for Model 3 including residual versus fitted value plots, Q-Q plots for checking the normality of residuals, Cook's distance for outliers, residual and high leverage point plots did not reveal substantial deviations from linear regression assumptions. Model 3 was thus selected as the best model. Reported results are thus based on this model.

>>>>

@@

\caption{\label{tab:tabmodel}Summary model performance measures.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & $\text{df}^a$ & $\text{AIC}^b$ & $\text{BIC}^c$ & $\text{MSE}^d$ & $R^2$ & Adjusted $R^2$\\
\midrule
Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
\addlinespace
Model 6 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
Model 7 & 41 & 15888.38 & 16117.55 & 173.69  & 0.36 & 0.34\\
Model 8 & 42 & 15849.19 & 16083.94 & 170.11  & 0.37 & 0.36\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
\end{tabular}
\end{table}
```
Likelihood ratio tests showed a statistically significant difference
between Model 1 and 2 with evidence of in favor of the latter
($\chi^2_{(1)} = 7064; p>0.05$). Model 2 which adjusted for all
potentially confounding covariates in \@ref(tab:tabthree) including an
interaction between cholestrol level and age was then compared to Model
3 which adjusted for only covariates we found significant in univariate
analyses. Results here similarly showed evidence in favor of Model 2
($\chi^2_{(1)} = 7979; p>0.05$). Comparing Model 2 to Model 3 revealed a
non-statistically difference in model fit suggesting that Model 3, even
though smaller performed comparably well. Model 3 was thus selected and
compared to Model 4, 5, 6, and 7 with evidence based on the LRT favoring
its choice as the best model ($p>0.05$). Additional regression
diagnostics for Model 3 including residual versus fitted value plots,
Q-Q plots for checking the normality of residuals, Cook's distance for
outliers, residual and high leverage point plots did not reveal
substantial deviations from linear regression assumptions. Model 3 was
thus selected as the best model. Reported results are thus based on this
model.

```{r tabfour, echo=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "MODEL 5: Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)
